---
title: "R_Data_Visualization_HW_Solutions"
author: "NYC Data Science Academy"
output: html_document
---
## Question #1: Scatterplot	

The data frame cars in the datasets package records the speed (in mph) and stopping distance (in ft) for 50 cars. Load the dataset using `data(cars)`

1. Create a scatterplot of `dist` (y-axis) vs. `speed` (x-axis).
2. Refine the basic plot by labeling the x-axis with "Speed (mpg)" and the y-axis with "Stopping Distance (ft)". Also add a title to the plot.
3. Revise the plot by changing the every point from the default open circles to red filled triangles (`col="red", pch=17`).

```{r cars, warning=FALSE, message = FALSE}

# Question 1

library(ggplot2)
data(cars)
plot(cars$speed,cars$dist)

ggplot(data = cars , aes(x = speed, y=dist))+
  geom_point(col="red",pch = 17)+
  xlab("Speed (mpg)")+
  ylab("Stopping Distance (ft)")+
  ggtitle("Relationship between Speed and Stopping Distance ")


```

## Question 2: Boxplot and Density Curves

The dataset `faithful` contains the duration of the eruptions (in minutes) and the waiting time before that eruption (in minutes) for the Old Faithful geyser. Load the dataset using `data(faithful)`.

1. In the faithful data frame, add a variable `length` that is "short" if the eruption is less than 3.2 minutes, and "long" otherwise.
2. Create parallel boxplots of the waiting times for the "short" and "long" eruptions.
3. Create overlapping density curves of the waiting times of the "short" and "long" eruptions.
4. Briefly describe your findings from the boxplots and the density curves.


```{r faithful}

# Question 2
data(faithful)

# 2.1
faithful$length = ifelse(faithful$eruptions < 3.2,
                         "short", "long")
faithful$length = as.factor(faithful$length)
class(faithful$length)

# 2.2
ggplot(faithful, aes(x = length, y = waiting)) +
  geom_boxplot()

# 2.3
ggplot(faithful, aes(x = waiting))+
  geom_density(aes(color = length))

# 2.4
# It seems that shorter wait times correspond to shorter eruption times
# Maybe if there is less waiting time, there is less build up, # and thus the eruption is shorter. 

```

## Question 3: Bar Plot and Histograms

Load the New York Knicks dataset, `Knicks.rda`. You will see a dataframe called `data` in your environment.

1. Calculate the winning ratio of New York Knicks in different seasons. Visualize how the winning ratio changes every year. (A bar plot is the most appropriate here.)
2. Calculate the winning ratio for both home and away games. (The row labelled with `visiting = 1` is an away game.) Create a bar plot to show home and away winning ratios for each season.
3. Plot five histograms to display the distribution of points in each season.
4. (**Challenge**) Calculate the average winning ratio and the average point-difference (i.e., `points-opp`) by each opponent. Create a scatter-plot to show winning ratio versus average point-difference. What pattern do you see in the graph?


```{r NBA}

# Question 3
library(dplyr)
load(url('https://s3.amazonaws.com/graderdata/Knicks.rda'))
knicks <- tbl_df(data)

# 3.1
knicks1 <- group_by(knicks, season) %>%
  summarise(ratio=sum(win=="W")/n())

ggplot(knicks1, aes(x=season, y=ratio)) + geom_bar(stat="identity")

# 3.2
knicks2 <- group_by(knicks, season, visiting) %>%
  summarise(ratio=sum(win=="W")/n())

ggplot(knicks2, aes(x=season,y=ratio)) +
  geom_bar(aes(fill=visiting), stat='identity',position='dodge')

# 3.3
ggplot(knicks,aes(x=points)) +
  geom_histogram()+
  facet_wrap(~season)

#### Difference between facet_wrap and facet_grid 
#### https://www3.nd.edu/~steve/computing_with_data/13_Facets/facets.html

# 3.4
knicks3 <- group_by(knicks, opponent) %>%
  summarise(ratio=sum(win=="W")/n(), diff=mean(points-opp))

ggplot(knicks3,aes(x=diff, y=ratio)) +
  geom_point(color='red4',size=4)+
  geom_hline(yintercept=0.5,colour='grey20',size=0.5,linetype=2)+
  geom_vline(xintercept=0,colour='grey20',size=0.5,linetype=2)+
  geom_text(aes(x=diff,y=ratio,label=opponent),
            hjust=0.7, vjust=1.4,angle = -30)+
  theme_bw()

# The plot shows that the more favorable the point differece, the higher the  win ratio, which sounds reasonable

```

## Question 4: Baby Names

Download the names.zip file from [here](https://s3.amazonaws.com/graderdata/names.zip). In the folder, you will find national data on the relative frequency of given names in the population of U.S. births where the individual has a social security number. For each year of birth YYYY after 1879, a comma-delimited file was created with a name format "yobYYYY.txt". Each record in the individual annual files has the format `name, sex, number` where `name` is 2 to 15 characters, `sex` is M (male) or F (female) and `number` is the number of occurrences of the name. Respond the following questions using the files in this database:

1. Write an R script to find how many unique names starting with "G" were given to babies in 2014.
2. How many babies were given a name starting with "K" in 2014?
3. How many names were given to both female and male babies in 2014? (i.e., The name "Jordan" counts because it was given to 6810 males and 1300 females that year.)
4. Write an R script to display the names with frequency between 5000 and 6000 in 2014. (*Hint: Be sure to first add the frequencies together for any name given to both males and females*.)
5. Create a bar plot showing the number of babies with names starting with A,B,C,D, ... Z for 2014. 
6. (**Challenge**) Create a line plot showing the change across years in proportion of babies with the names "Chris", "Christian", and "Christopher".
7. (**Challenge**) Create a line plot showing the change across years in number of male and female babies. 


```{r babyNames}

# Question 4
# 4.1
library(dplyr)
yob2014<- read.csv("names/yob2014.txt", 
                     header = FALSE,
                     col.names = c("name","sex","number"),
                     stringsAsFactors = FALSE)
yob2014$initial<- substring(yob2014$name, first = 1, last = 1) 
#first tells you to start by grabbing the first character, 
#last tells you to stop after grabbing the first character

subset <- dplyr::filter(yob2014, initial=="G") 
nlevels(as.factor(subset$name))

# 4.2
subset <- filter(yob2014, initial=="K")
sum(subset$number)

# 4.3
male <- filter(yob2014, sex=="M")
female <- filter(yob2014, sex=="F")
nrow(inner_join(male, female, by="name"))

# 4.4
group_by(yob2014, name) %>% 
  summarise(total=sum(number)) %>%
  filter(total >= 5000 & total <= 6000)
                
# 4.5
names <- group_by(yob2014, initial) %>%
        summarise(total = sum(number))

ggplot(names) + 
  geom_bar(aes(x=initial, y=total), stat='identity')

# 4.6
# Get the files names
setwd('names')
files = list.files(pattern="*.txt")
# First apply read.csv, then rbind
handle_each <- function(x){
  year <- as.numeric(gsub("[^\\d]+", "", x, perl=TRUE))
  temp <- read.csv(x, header = FALSE,  col.names = c("name","sex","number"))
  temp$year <- year
  return(temp)
}
myfiles = do.call(rbind, lapply(files, handle_each))
ccc = myfiles[with(myfiles, name %in% c('Chris', 'Christian', 'Christopher')), ]

total = group_by(myfiles, year) %>% summarise(total_number=sum(number))
sub = group_by(ccc, name, year) %>% summarise(number=sum(number))
result = left_join(sub, total, by="year") %>% mutate(ratio=number/total_number)

ggplot(data=result, aes(x=year, y=ratio)) +
  geom_line(aes(color=name))

# 4.7
sex = group_by(myfiles, year, sex) %>% summarise(number=sum(number))
ggplot(data=sex, aes(x=year, y=number)) +
  geom_line(aes(color=sex))

```


